---
title: "figure7g_size_distri"
output:
  pdf_document: default
  html_document: default
date: "2025-01-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Compare the distribution of size, figur 7g
This scripts compare the distribution of sphere size in mTOR and 5XF group. To generate figure 7g, you only need to run the sections "Statistics test" and "QQ plot"

There are two experimental groups, each has several animals. 

First, prepare the dataset
```{r}
library(ggplot2)
library(readxl)
library("viridis") 
library("ggpubr")


# Load data
file_path <- "./data/fig7g_statistics.xlsx"
data <- read_excel(file_path, sheet = "Fig7G_area continue")

# Clean data by removing rows with NaN
id_var <- rownames(data)
#col_names <- col_names
long_data <- reshape2::melt(data,variable.name = "Condition", value.name = "Value")

# Remove rows with NaN values in the reshaped data
long_data <- na.omit(long_data)

# Combine all data into a single group for overall density
long_data$Condition <- as.factor(long_data$Condition)

# Group conditions by prefix ("mTOR" or "5XF")
long_data$Group <- ifelse(grepl("^mTOR", long_data$Condition), "mTOR Group", 
                          ifelse(grepl("^5XF", long_data$Condition), "5XF Group", "Other"))

# Plot probability density for each condition and the combined data
ggplot(long_data, aes(x = Value, color = Condition)) +
  geom_density() +
  geom_density(data = long_data, aes(x = Value), inherit.aes = FALSE, 
               color = "black", linetype = "dashed") +
  labs(title = "Probability Density under Each Condition", 
       x = "Value", y = "Density", color = "Condition") +
  theme_minimal()
```


```{r}
# Separate data by groups
mtor_data <- subset(long_data, Group == "mTOR Group")
f5x_data <- subset(long_data, Group == "5XF Group")

# Plot PDF
pdf_plot <- ggplot() +
  # Individual mTOR experiment curves (light gray)
  geom_density(data = mtor_data, aes(x = Value, group = Condition), color = "lightgray", size = 0.5) +
  # Overall mTOR group curve (black)
  geom_density(data = mtor_data, aes(x = Value), color = "black", size = 1.2) +
  # Individual 5XF experiment curves (light blue)
  geom_density(data = f5x_data, aes(x = Value, group = Condition), color = "lightblue", size = 0.5) +
  # Overall 5XF group curve (dark blue)
  geom_density(data = f5x_data, aes(x = Value), color = "blue", size = 1.2) +
  theme_pubr() +
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )
ggsave("pdf_mTOR_5XF.pdf",dpi = 600, width = 4.2, height = 3.5)
#print(pdf_plot)
```

```{r}
# Plot PDF
pdf_plot2 <- ggplot() +
  # Individual mTOR experiment curves (light gray)
  geom_density(data = mtor_data, aes(x = Value, group = Condition), color = "lightgray", size = 0.5) +
  # Overall mTOR group curve (black)
  geom_density(data = mtor_data, aes(x = Value), color = "black", size = 1.2) +
  # Individual 5XF experiment curves (light blue)
  geom_density(data = f5x_data, aes(x = Value, group = Condition), color = "lightblue", size = 0.5) +
  # Overall 5XF group curve (dark blue)
  geom_density(data = f5x_data, aes(x = Value), color = "darkblue", size = 1.2) +
  # Add legend manually
  scale_color_manual(name = "Group", values = c("mTOR" = "black", "5XF" = "darkblue"), 
                     labels = c("mTOR Group", "5XF Group")) +
  theme_pubr()
  ggsave("mTOR_5XF_PDF_plot.pdf", plot = pdf_plot2, width = 8, height = 6)
  print(pdf_plot2)
```


## Group the data by experimental conditions
Conditions with mTOR and condition with 5XFAD

```{r}
# Group conditions by prefix ("mTOR" or "5XF")
long_data$Group <- ifelse(grepl("^mTOR", long_data$Condition), "mTOR Group", 
                          ifelse(grepl("^5XF", long_data$Condition), "5XF Group", "Other"))

# Plot probability density for grouped conditions
# ggplot(long_data, aes(x = Value, color = Condition)) +
#   geom_density() +
#   facet_wrap(~ Group, scales = "free_y") +
#   labs(title = "Probability Density by Grouped Conditions", 
#        x = "Value", y = "Density", color = "Condition") +
#   theme_minimal()


# lump data togeter
ggplot(long_data, aes(x = Value, color = Group)) +
  geom_density() +
  labs(title = "Probability Density", 
       x = "Value", y = "Density", color = "Condition") +
  theme_minimal()

```
### ECDF
plot and compare the emperical cumulative density functions

```{r}
# Plot empirical cumulative density function (ECDF) for grouped conditions
# ggplot(long_data, aes(x = Value, color = Condition)) +
#   stat_ecdf() +
#   facet_wrap(~ Group, scales = "free_y") +
#   labs(title = "Empirical Cumulative Density Function by Grouped Conditions", 
#        x = "Value", y = "ECDF", color = "Condition") +
#   theme_minimal()


ggplot(long_data, aes(x = Value, color = Group)) +
  stat_ecdf() +
  labs(title = "Empirical Cumulative Density Function", 
       x = "Value", y = "ECDF", color = "Group") +
  theme_minimal()
```

## Statistics test
Show that the 5XF group has more fraction of larger size than 10 mu^2. We use Two-proportion z-test
```{r}
thresold <- 10
mtor_data <- subset(long_data, Group == "mTOR Group")
f5x_data <- subset(long_data, Group == "5XF Group")

# Calculate proportions
mtor_proportion <- mean(mtor_data$Value > thresold)
f5x_proportion <- mean(f5x_data$Value > thresold)

# Count successes and totals for z-test
mtor_success <- sum(mtor_data$Value > thresold)
mtor_total <- nrow(mtor_data)

f5x_success <- sum(f5x_data$Value > thresold)
f5x_total <- nrow(f5x_data)

# Perform two-proportion z-test
prop_test <- prop.test(c(f5x_success, mtor_success), c(f5x_total, mtor_total), alternative = "greater")

# Print results
cat("Proportion of values > 10:\n")
cat("mTOR Group: ", mtor_proportion, "\n")
cat("5XF Group: ", f5x_proportion, "\n\n")
cat("Two-proportion z-test results:\n")
print(prop_test)

# Calculate fraction of values > 10 for each experimental condition
fraction_by_condition <- aggregate(Value > thresold ~ Condition, data = long_data, mean)
colnames(fraction_by_condition) <- c("Condition", "Fraction > 10")

# Print fraction results
cat("\nFraction of values > 10 by experimental condition:\n")
print(fraction_by_condition)
```


### QQ plot to compare the two ECDF
Use emperical QQ plot

```{r}
my.theme = list(theme(axis.line.y.left = element_line(linewidth = 0.5, color = "black"),
                      axis.line.y.right = element_line(linewidth = 0.5, color = "black"),
                      axis.line.x.top =  element_line(linewidth = 0.5, color ="black"),
                      axis.line.x.bottom = element_line(linewidth = 0.5, color = "black")))

qq.out <- qqplot(x=mtor_data$Value, y=f5x_data$Value, plot.it=FALSE)
qq.out <- as.data.frame(qq.out)

names(qq.out) <- c("mTOR", "AD")  # Add batch names to output
# Set the x and y limits
xylim <- range( c(qq.out$mTOR, qq.out$AD) )

gp <- ggscatter(qq.out, x = "mTOR", y = "AD",
                size = 1.5, alpha = 0.5
          #add = "reg.line",                                 # Add regression line
          #conf.int = TRUE,                                  # Add confidence interval
          ) + xlim(0, 65) + ylim(0, 65) + coord_fixed() +
  geom_abline(intercept = 0, slope = 1) + 
  geom_hline(yintercept = 10,linetype = "dashed", size = 0.2) + 
  geom_vline(xintercept = 10,linetype = "dashed", size = 0.2) +
  font("xlab", size = 12) + font("ylab", size = 12) + 
  my.theme
ggsave("qqplot_mTOR_5XF.pdf",dpi = 600, width = 3.5, height = 3.5)
```


Using slight fewer quantiles to avoid clutter, we also color the qunatiles
```{r}
# library("viridis") 
# library(RColorBrewer)

n = max(length(mtor_data$Value), length(f5x_data$Value))
p = (1:n - 1)/(n - 1)
# p = (1:1000)/1000
q_m = quantile(mtor_data$Value,p)
q_f = quantile(f5x_data$Value,p)
qs = data.frame(mTOR = q_m, F5X = q_f, Quantile = p*100)
#plot(q_m, q_f)
new_qq <- ggplot(qs,aes(x = mTOR, y= F5X, colour = Quantile)) + geom_point(size = 1.5) + 
  coord_fixed() +
  xlim(0, 65) + ylim(0, 65)  +
  scale_color_viridis(option = "D") +
  # scale_fill_distiller(palette = "Spectral") + 
  geom_abline(intercept = 0, slope = 1, size = 0.5, colour = "gray") +
  geom_hline(yintercept = 10,linetype = "dashed", size = 0.2) +
  geom_vline(xintercept = 10,linetype = "dashed", size = 0.2) +
  font("xlab", size = 12) + font("ylab", size = 12) +
  theme_classic() + 
  theme(axis.title.y = element_text(size = 12),
        axis.title.x = element_text(size = 12),
        axis.text = element_text(size = 10),
        legend.position = "top")

# save as a pdf
ggsave("qq_plot_colored.pdf",dpi = 300, width = 4, height = 3.5)
# new_qq

# make a scatter plot color the dot with values
```

### lognormal distribution
The distribution of size can be well captured by an lognormal distribution

```{r}
# Plot PDF
mtor_log <- mtor_data
mtor_log$Value <- log10(mtor_data$Value)
f5x_log <- f5x_data
f5x_log$Value <- log10(f5x_data$Value)

pdf_plot <- ggplot() +
  # Individual mTOR experiment curves (light gray)
  geom_density(data = mtor_log, aes(x = Value, group = Condition), color = "lightgray", size = 0.5) +
  # Overall mTOR group curve (black)
  geom_density(data = mtor_log, aes(x = Value), color = "black", size = 1.2) +
  # Individual 5XF experiment curves (light blue)
  geom_density(data = f5x_log, aes(x = Value, group = Condition), color = "lightblue", size = 0.5) +
  # Overall 5XF group curve (dark blue)
  geom_density(data = f5x_log, aes(x = Value), color = "blue", size = 1.2) +
  xlab("log10(Size)") + ylab("pdf") + 
  theme_pubr() +
  theme(
    legend.position = "none",
    plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10)
  )
ggsave("lognorma_pdf.pdf",dpi = 600, width = 4, height = 3)
```
```{r}
# Plot empirical cumulative density function (ECDF) for grouped conditions
long_log <- long_data
long_log$Value <- log10(long_data$Value)
ggplot(long_log, aes(x = Value, color = Condition)) +
  stat_ecdf() +
  facet_wrap(~ Group, scales = "free_y") +
  labs(title = "Empirical Cumulative Density Function by Grouped Conditions", 
       x = "Value", y = "ECDF", color = "Condition") +
  theme_minimal()
ggsave("ecdf_log_size_compare.pdf", dpi = 300, width = 8, height = 4)
```
```{r}
require(qqplotr)

customization <- list(scale_fill_brewer(palette = "Set2"),
                      scale_colour_brewer(palette = "Set2"),
                      xlab("Normal quantiles"),
                      ylab("Height (in.)"),
                      coord_equal(),
                      theme_light(),
                      theme(legend.position = c(0.8, 0.2), aspect.ratio = 1))

mtor_data %>%
  #ggplot(aes(sample = Value, colour = Gender, fill = Gender)) +
  ggplot(aes(sample = Value)) +
  geom_abline(colour = "grey40") +
  stat_qq_band(bandType = "ks", distribution = "lnorm", alpha = 0.4) +
  stat_qq_line(distribution = "lnorm") +
  stat_qq_point(distribution = "lnorm") +
  customization +
  #facet_grid(. ~ Gender) +
  xlab("Lognormal quantiles") +
  ylab("Weight (kg.)") +
  theme(legend.position = c(0.9, 0.2))
```


```{r}
f5x_data %>%
  #ggplot(aes(sample = Value, colour = Gender, fill = Gender)) +
  ggplot(aes(sample = Value)) +
  geom_abline(colour = "grey40") +
  stat_qq_band(bandType = "ks", distribution = "lnorm", alpha = 0.4) +
  stat_qq_line(distribution = "lnorm") +
  stat_qq_point(distribution = "lnorm") +
  customization +
  #facet_grid(. ~ Gender) +
  xlab("Lognormal quantiles") +
  ylab("Weight (kg.)") +
  theme(legend.position = c(0.9, 0.2))
```


```{r}
long_log %>%
  ggplot(aes(sample = Value, colour=Group, fill=Group)) +
  geom_abline(colour = "grey40") +
  stat_qq_band(bandType = "ts", alpha = 0.4) +
  stat_qq_line() +
  stat_qq_point() +
  customization +
  ylab("log_10 Size (um^2)") 
  
  ggsave("qqplot_lognorm_compare_mTOR_5XF.pdf",dpi = 600, width = 3.5, height = 3.5)
```